\documentclass[thesis.tex]{subfiles} 
\begin{document}

\section{Before opinion polls}

\subsection{Fundamental determinants of voter preference}

The models and techniques examined in this paper rely primarily on polling data for an up-to-date estimate of popular opinion. But even without opinion polling, presidential elections can be predicted ahead of time using various indicators.

For example, \cite{Hibbs:2008aa} presents a Bread-and-Peace model of presidential elections, where he attempts to explain persistent fundamental determinants of election outcomes in a linear regression model. He looks at two variables, \begin{enumerate}
	\item the weighted-average growth of per-capita real personal disposable income over the previous term;
	\item and U.S. military fatalities in unprovoked, hostile deployments of American forces abroad,
\end{enumerate} and interpets elections as mostly referendums on the White House party's economic record, along with a substantial aversion to foreign wars. 

States popular votes can also be estimated ahead of time, using similar indicators such as economic factors, and further intuitive ones such as general political leaning or home-state advantage (\citealt{Campbell:1992aa,Campbell:2006aa}).

The models analyzed in this paper use these estimates as prior knowledge on election outcomes, to be incorporated with polling data.

\subsection{State deviance between election cycles}

For states, we can note that deviations from the national average are relatively consistent between election cycles \cite{Lock:2010aa}. This means that a state that leans more Democratic than the nation during one election, will probably lean the same way during the next election.

\cite{Lock:2010aa} handle this using a multilevel model, to compensate for the small number of data points for each state (looking at the last eight election cycles gives only eight data points for each state). Multilevel modeling takes advantage of both within-group and between-group information. Groups with poor information will get pulled towards the mean, and groups with rich information will weight their own observations more highly. The technique is also used in education, demographics, and geographical data \citep{Ghitza:2013aa,Gelman:2006aa,Aitkin:1981aa}.

In this case, we want to estimate variance in state deviations from the national vote between election cycles, \[
  \frac{1}{N}\sum_{y=1}^N (d_{s, y} - d_{s, y - 1})^2,
\] where \(N\) denotes the total number of election cycles we're looking at and \(d_{s, y}\) denotes the deviation from the national vote in election cycle \(y\) for state \(s\).

The multilevel model, then, looks like \begin{equation*}
	\left[(d_{s, y} - d_{s, y - 1})^2\right]_i \sim \mu + \gamma_s + \epsilon_i, \quad
	\gamma_s \sim \N(0, \sigma_\gamma^2), \quad
	\epsilon_i \sim \N(0, \sigma_\epsilon^2),
\end{equation*} where \(\mu + \gamma_s\) will give us the estimated variance in deviations for state \(s\). We describe this class of model further in Section 6.

% TODO: Graph?

\begin{comment}

For states, we can also note that deviations from the national average are relatively consistent between election cycles \cite{Lock:2010aa}. Here's a graph maybe. This means that a state that leans more Democratic than the nation during one election, will probably lean the same way during the next election. The model looks sort of like: \[
	a model,
\] where the things mean this. This gives us the mean for each state, this is how you put it into \texttt{lmer}.

We handle this using multilevel modeling, as in \cite{Lock:2010aa}. \textbf{What is multi-level regression?} Also called mixed-effects (it's like balancing between no effects and fixed effects). Takes advantage of both within-group and between-group information. If a group has very little data (like, state deviations from national vote over the last seven elections), it's ``shrunk'' towards the overall mean (shrinkage estimation). Also used in education, demographics, and geographical data \citep{Ghitza:2013aa,Gelman:2006aa,Aitkin:1981aa}.

The general ideas behind multilevel modeling were described in \cite{Harville:1977aa}, we see some equivalent approaches in \cite{Goldstein:1986aa} and \cite{Goldstein:1989aa}, and the modern approach using REML and Expectation-Maximization is described through example in \cite{Aitkin:1981aa}. \cite{Gelman:2006aa} gives a recent example of using and interpreting the results of multilevel modeling, using the \texttt{lme4} \citep{lme4} package in R \citep{R}. For more information, see \cite{Gelman:2007aa}.

Some people also solve it using the Bayesian approach \citep{Gelman:2003aa,Price:1996aa}, but we won't use that.

\end{comment}

\end{document}